---
title: "Effect of Population Density on Life Expectancy"
author: "Maria Benavides"
date: "`r lubridate::today()`"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(geonames)
library(countrycode)
library(here)
library(broom)
library(kableExtra)

key <- getOption("geonamesUsername") # Call geonames username to access data
theme_set(new = theme_minimal())
```

```{r import & merge data}
# Import data 
gapminder_df <- gapminder::gapminder
geonames_df <- GNcountryInfo() %>%
  rename(
    country = countryCode
  )

# Create key to join
gapminder_key <- gapminder_df %>%
  mutate(
    country = countrycode(
      country, 
      origin = "country.name", 
      destination = "iso2c"
    )
  )

# Merge dataframes based on the country variable
all_variables <- left_join(
  gapminder_key, geonames_df, by = "country"
  ) %>%
  select( # Select only variables of interest for this exercise
   country,
   lifeExp,
   pop,
   areaInSqKm, 
   year
  ) %>%
  mutate(
    areaInSqKm = as.numeric(areaInSqKm))

# Add new variable for pop density 
all_variables <- all_variables %>%
  mutate(
    pop_density = pop / areaInSqKm
  )


```

```{r correlation lifeExp pop density}
# Run a regression model 
regression_popden_lifeexp <- lm(lifeExp ~ pop_density, data = all_variables)
regr_table <- tidy(regression_popden_lifeexp) %>%
  rename(
    "Predictor" = term,
    "B" = estimate, 
    "SE" = std.error, 
    "t" = statistic, 
    "p" = p.value
    ) 
kable(regr_table, 
      digits = 3, 
      caption = "Table correlation between life expectancy and population density"
      ) 
```

Using a simple OLS model that considers life expectancy as the dependent variable and population density as the independent, we see that there is a positive correlation between those, which is statistically significant; this suggests that places with higher population density have a higher life expectancy. 

```{r correlation graph}
#Plot the correlation
ggplot(all_variables, aes(x=pop_density, y=lifeExp)) + 
  geom_point(alpha = 0.2) +
  geom_smooth(method=lm)+
  labs(
    title = "Correlation between life expectancy and population density", 
    x = "Population density", 
    y = "Life expectancy"
  )

```

The graph is shows visually what the model predicted, that is to say, there is a positive significant correlation between life expectancy and population density. Nonetheless, it is worth mentioning that this graph is showing us that there are outliers in our data frame. Singapore and Hong Kong are extremely dense, and might biases these results. 

## Session info

```{r, echo = TRUE}
devtools::session_info()
```
